Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-18 analysis/index.Rmd
Remove trailing space
Add newer abstract

Over the past two decades, international non-governmental organizations (INGOs) have faced increasing legal scrutiny and repression abroad. As a result, official aid to INGOs in repressive countries has reduced. NGOs thus need access to a diverse array of funds and may need to increase fundraising from private donors. Which demographic and experiential factors influence the decision of individuals to donate to legally besieged NGOs? We investigate this question using a market simulation based on a nationally representative conjoint experiment conducted on individuals with a history of charitable giving. We find that trust in political institutions and charities is enormously influential in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to human rights INGOs with friendly relationships with host governments, and that support hold steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. Our findings have practical implications for INGOs that aim to raise funds from the general public amid closing civic space.

>>>>

The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals' preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
